Interisland movements and oceanic swimming of woodland caribou in Newfoundland

### 2020-03-18 paper/manuscript.Rmd
edit wd - thanks alec for all the help it's been really great thanks for being so helpful
add number

```{r input}
edges <- readRDS('output/island-edges.Rds')
icefree <- edges[season == 'icefree']
```

@@

@@

The Fogo Island archipelago, off the coast of Newfoundland, Canada, is home to
approximately 300 woodland caribou. During routine fieldwork on May 30, 2017, we
observed an unmarked adult male caribou swim between Western and Eastern Indian
Islands (Figure 1), a distance of at least 470 m which took ~9 minutes (~52 m
per minute). This observation prompted us to investigate whether swimming
behaviour was common among GPS radio-collared caribou in the population (n = 29
adult females; 1240g, GPS 4400M; Lotek Wireless Inc., Newmarket, Ontario,
Canada). We identified swimming events as two consecutive GPS (2-hour relocation
rates) locations from an individual occurred on different islands. Northeastern
Newfoundland typically experiences pack ice during winter and caribou are known
locally to travel between islands by walking over the ice. We restricted our GPS
data to the ice-free period of the year (April 1 - December 31) and we
identified  adult female caribou that collectively had ``` `r nids` swimming events
over three years (Figure 2). In addition to our own visual and remotely sensed
observations, residents of Fogo Island have also reported observing caribou
swimming between islands on numerous occasions.

<!-- add numbers in here -->
In total, `r nevents` swimming events occurred throughout the ice-free season, with
a peak in November. On average, caribou swam between islands approximately ever
XX days, but some individuals swam up to XX times per year (range XX-XX).
Caribou may swim between islands in the ocean for similar reasons they swim in
freshwater [@Leblond2016]. We propose that one such explanation, the forage
limitation hypothesis, is a likely explanation for oceanic swimming for caribou
that live on islands in the Fogo Island archipelago.
<!--Is it better
to just state our results here, or give this sort of teaser? Initially I had "86
events by 3 individuals" here, but then we ignore that for a couple paragraphs
of context before getting back into it... which seemed weird
QW: I agree. I've added some Fogo context-->

>>>>

```{r input}
edges <- readRDS('../output/island-edges.Rds')
icefree <- edges[season == 'icefree']
```

@@

# Max per individual
neventidMax <- max(neventid$N)

# Mean per individual
neventidMean <- round(mean(neventid$N))

# Min per individual
neventidMin <- min(neventid$N)

# Duration of time per island
dur <- round(icefree[, mean(islandlen)], digits = 0)
durMin <- round(icefree[, min(islandlen)], digits = 0)
durMax <- round(icefree[, max(islandlen)], digits = 0)

@@

The Fogo Island archipelago, off the coast of Newfoundland, Canada, is home to
approximately 300 woodland caribou. During routine fieldwork on May 30, 2017, we
observed an unmarked adult male caribou swim between Western and Eastern Indian
Islands (Figure 1), a distance of at least 470 m which took ~9 minutes (~52 m
per minute). This observation prompted us to investigate whether swimming
behaviour was common among GPS radio-collared caribou in the population (n = 29
adult females; 1240g, GPS 4400M; Lotek Wireless Inc., Newmarket, Ontario,
Canada). We identified swimming events as two consecutive GPS (2-hour relocation
rates) locations from an individual occurred on different islands. Northeastern
Newfoundland typically experiences pack ice during winter and caribou are known
locally to travel between islands by walking over the ice. We restricted our GPS
data to the ice-free period of the year (April 1 - December 31) and we
identified  adult female caribou that collectively had `r nids` swimming events
over three years (Figure 2). In addition to our own visual and remotely sensed
observations, residents of Fogo Island have also reported observing caribou
swimming between islands on numerous occasions.

<!-- add numbers in here -->
In total, `r nevents` swimming events occurred throughout the ice-free season,
with a peak in November. On average, caribou swam approximately `r neventidMean`
(range = `r neventidMin`-`r neventidMax`) per year and remained on those islands
for approximately `r dur` (range = `r durMin`-`r durMax`) days before swimming again. Caribou may swim between
islands in the ocean for similar reasons they swim in freshwater [@Leblond2016].
We propose that one such explanation, the forage limitation hypothesis, is a
likely explanation for oceanic swimming for caribou that live on islands in the
Fogo Island archipelago.
<!--Is it better
to just state our results here, or give this sort of teaser? Initially I had "86
events by 3 individuals" here, but then we ignore that for a couple paragraphs
of context before getting back into it... which seemed weird
QW: I agree. I've added some Fogo context-->
